---
title: "Regular expressions in R - solutions to in-class practice"
date: "`r lubridate::today()`"
output: html_document
always_allow_html: true
---

```{r}
library(stringr)
suppressPackageStartupMessages(library(tidyverse))
```


**PRACTICE 1.** Use dots and `$` to match the last four characters of all sentences in `x`: 

```{r}
x <- c("This is a sentence to practice regular expressions", 
       "Find the word regex in this sentence",
       "A last sentence similar to the first sentence")
```

```{r practice1-solution}
str_view(x, "....$")
```


**PRACTICE 2.** This sentence contains a set of ugly characters. Identify them first using the `|` operator, then using the `[]`:

```{r}
j <- "This&is#MACS%30500"
```

```{r practice2-solution}
str_view_all(j, "&|#|%")
str_view_all(j, "[&#%]")
```


**PRACTICE 3.** Apply these meta-characters to extract all phone numbers from the following string:

```{r}
s <- "My number is 444-3340, Carl number is 555-3755, Richard number is 666-6681"
```

```{r pratice3-solution}
str_view_all(s, "[0-9][0-9][0-9]-[0-9][0-9][0-9][0-9]")
str_view_all(s, "\\d\\d\\d-\\d\\d\\d\\d")
str_view_all(s, r"(\d\d\d-\d\d\d\d)")
```


**PRACTICE 4.** Match all three emails in the following string:

```{r}
email <- "smith90@uchicago.edu nardin@uchicago.edu happy2@uchicago.edu"
```

```{r pratice4-solution}
str_view_all(email, "[a-z]+[0-9]*@uchicago\\.edu")
```

Unpack the solution for practice 4: `[a-z]+` matches one or more characters, `[0-9]*` matches zero or more numbers, `@uchicago\.edu` is a literal match with `\` to match the dot literally
