---
title: "Text analysis with R tidyverse: Jane Austen example"
output: github_document
---


Tutorials taken from 
* [Chapter 1.3 Tidying the works of Jane Austen](https://www.tidytextmining.com/tidytext.html#tidyausten)
* [Chapter 3.1 Term frequency in Jane Austen’s novels](https://www.tidytextmining.com/tfidf#term-frequency-in-jane-austens-novels)
* [Chapter 4 Relationships between words: n-grams and correlations](https://www.tidytextmining.com/ngrams)


Import libraries 
```{r}
library(tidyverse)
library(tidytext)
library(janeaustenr)
library(dplyr)
library(stringr)
library(ggplot2)
library(forcats)
```

The `janeaustenr` package provides Jane Austen's texts in a one-row-per-line format, where a line is analogous to a literal printed line in a physical book. 

Explore data (see See https://github.com/juliasilge/janeaustenr)
```{r}
austen_books()

is.vector(emma)
emma[1:20]

```

Get data and perform basic manipulation (e.g., use `mutate()` to annotate a linenumber and chapter)
```{r}
original_books <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, 
                                     regex("^chapter [\\divxlc]",
                                           ignore_case = TRUE)))) %>%
  ungroup()
```

What does this regular expression mean? Break it down: `^chapter`, `\\`, `[divxlc]`
It matches thw word chapter lowercase, and any of the literal roman characters that follow it 

Pre-process data by reshaping this df into the one-token-per-row format using `unnest_tokens`
```{r}
tidy_books <- original_books %>%
  unnest_tokens(word, text)

head(tidy_books)
tail(tidy_books)
```

Pre-process data by removing stop words
```{r}
data(stop_words)
```

Notice `anti_join()` return all rows from x without a match in y 
See docs for more: https://dplyr.tidyverse.org/reference/filter-joins.html especially if oyu need to specify column names on the two dataframes using by
```{r}
tidy_books_no_stop <- tidy_books %>%
  anti_join(stop_words)
```

Remove stop word using simple `filter()`
```{r}
tidy_books  %>% 
 filter(!word %in% stop_words$word)
```


Data analysis: use count() to find most common words in all books
```{r}
tidy_books %>%
  count(word, sort = TRUE)
```


Data analysis: filter, reorder, and visualize common words
```{r}
tidy_books %>%
  count(word, sort = TRUE) %>%
  filter(n > 600) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
  geom_col() +
  labs(y = NULL)

```


Data analysis (CHAPTER 3): TF-IDF 

We start with TF or Term Frequency: 
TF is simply the count of how frequent a word occurs in a document. Here our documents are books. If we want to know the TF of a given word of interest by book, we need: 
the count of all words, the count of all words by book, we take the word of interest and divide by the total words in that wook (word/words)

```{r}
# count all words
book_words <- austen_books() %>%
  unnest_tokens(word, text) %>%
  count(book, word, sort = TRUE)
book_words

```

```{r}
# summarize total words grouped by each book
total_words <- book_words %>% 
  group_by(book) %>% 
  summarize(total = sum(n))
total_words
```

```{r}
# join these data by book
book_words <- left_join(book_words, total_words)
book_words
```

This book_words data frame has one word for each word-book combination
* n: number of times that word is used in that book
* total: total words in that book
So we can plot our distribution of n/total

```{r}
# look at the n/total for each book
ggplot(book_words, aes(n/total, fill = book)) +
  geom_histogram(show.legend = FALSE) +
  xlim(NA, 0.0009) +
  facet_wrap(~book, ncol = 2, scales = "free_y")
```

Zipf's law: we notice many words occur rarely, and few words occur very frequently. In other terms the frequency that a word appears is inversely proportional to its rank: https://simple.wikipedia.org/wiki/Zipf%27s_law

See book for code and examples. 

Now we calculate TF-IDF
* TF is frequency count of words per document 
* IDF measures how important a word is to a document by:
  * decreasing the weight for commonly used words and 
  * increasing the weight for words that are rarely used in corpus of documents
For a theoretical explanation of tf-idf: https://web.stanford.edu/~jurafsky/slp3/6.pdf


To calculate TF-IDF in tidy R, use the `bind_tf_idf()` function, which takes a tidy text df as input with three columns: the single token (word here), the single document (book here), and the count that is hoe many time each single document contains each single token (n here)

```{r}
book_tf_idf <- book_words %>%
  bind_tf_idf(word, book, n)
book_tf_idf
```

Notice that idf and thus tf-idf are zero for the most common words. Now look at terms with high tf-idf by rearranging the previous results 

```{r}
book_tf_idf %>%
  select(-total) %>%
  arrange(desc(tf_idf))
```

Notice those are proper nouns important to each of these novels: none of them appear in all novels, and each is characteristic of a particular one

Visualize the results: 
```{r}
book_tf_idf %>%
  group_by(book) %>%
  slice_max(tf_idf, n = 15) %>%
  ungroup() %>%
  ggplot(aes(tf_idf, fct_reorder(word, tf_idf), fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free") +
  labs(x = "tf-idf", y = NULL)
```


Data analysis (CHAPTER 4): relationship between words (n-grams and correlations)
Follow the chapter and copy/paste the code to reproduce it!